---
title: "rRNA RibOxi-Seq summarization for L.major"
author: "Yinzhou Zhu"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
```

# In this sctipt, analysis is done with L.*major*  


## Import the counts from BED files  
```{r import counts, echo=FALSE}
files_to_read<- str_remove(dir(pattern="LM[1-9]\\.bed"), pattern = "\\.bed")

for (each_file in files_to_read){
  assign(each_file, read_delim(paste(each_file, ".bed", sep = ""), "\t", escape_double = FALSE, col_names = FALSE,
                     col_types = cols(X5 = col_skip(), X4 = col_skip()), trim_ws = TRUE))
  }
```


## Counting all 3' ends of the alignments  
```{r tally end counts, echo=FALSE}
for (each_df in ls(pattern = "LM.$")){
    assign(paste(each_df, "_3_end", sep = ""), get(each_df)%>%rename(gene_symbol=X1, start=X2, end=X3, strand=X6)%>%filter(grepl("LM",gene_symbol)&(strand=="+"))%>%group_by(gene_symbol, end)%>%tally(name = each_df))
}
full_list <- get(paste(each_df, "_3_end", sep = ""))[FALSE,]
colnames(full_list)<-c("gene_symbol", "end", "n")
```

## Output summarized counts  
```{r join samples and write to file}
for (each_file in files_to_read){
    full_list <- data.frame(full_join(full_list, get(paste(each_file, "_3_end", sep = ""))))
  }
full_list<-full_list%>%replace(is.na(.), 0)
full_list$n<-NULL
write.csv(full_list, file = "LM_counts.csv", row.names = FALSE, col.names = FALSE)
```

## Visualization  

### Raw counts  

These are the raw counts for all three replicates, and the black bars are std error bars  



```{r raw counts}
full_list%>%
        gather(key="Samples", value = "counts", starts_with("LM"))%>%
          ggplot(aes(x=end, y=counts))+geom_col(aes(fill=Samples)) + 
            xlab("Position") + ylab("Count") + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + facet_wrap(vars(gene_symbol), ncol = 1, scales = "free") + stat_summary(geom = "errorbar", fun.data = mean_sdl, position = "dodge")
```
\newpage  

### Scores  

For a initial look the calculation for scores is as follows:

1. Calculate mean for each position by averaging all 3 replicates (row means)  

2. Calculate mean for the means of all positions (column mean of row means)  

3. score = row means/(column mean of row means)  

Since the 3 samples are pretty balanced in terms of total reads, so inter-sample normalization was not done  



```{r normalized counts}
full_list%>%
        mutate(LM1=LM1/mean(LM1), LM4=LM4/mean(LM4),LM5=LM5/mean(LM5), .keep="unused") %>%
        transmute(gene_symbol, end, norm_score = rowMeans(select(., -c(gene_symbol,end)))) %>%
          ggplot(aes(x=end, y=norm_score))+geom_col() + 
            xlab("Position") + ylab("Scores") + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + facet_wrap(vars(gene_symbol), ncol = 1, scales = "free") + coord_cartesian(ylim = c(0, 10))

LM_score<-full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused")
write.csv(LM_score, file = "LM_scores.csv")
```
\newpage

## Fine tuning score cutoffs

**Score cutoff: 2**  

```{r normalized counts2}
full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused") %>%
        filter(norm_score>2) %>%
          ggplot(aes(x=end, y=norm_score))+geom_col() + 
            xlab("Position") + ylab("Scores") + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + facet_wrap(vars(gene_symbol), ncol = 1, scales = "free") + coord_cartesian(ylim = c(0, 15))

LM_score<-full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused")
```
\newpage
**Score cutoff: 1**  

```{r normalized counts3}
full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused") %>%
        filter(norm_score>1) %>%
          ggplot(aes(x=end, y=norm_score))+geom_col() + 
            xlab("Position") + ylab("Scores") + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + facet_wrap(vars(gene_symbol), ncol = 1, scales = "free") + coord_cartesian(ylim = c(0, 15))

LM_score<-full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused")
```
\newpage
**Score cutoff: 3**  

```{r normalized counts4}
full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused") %>%
        filter(norm_score>3) %>%
          ggplot(aes(x=end, y=norm_score))+geom_col() + 
            xlab("Position") + ylab("Scores") + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + facet_wrap(vars(gene_symbol), ncol = 1, scales = "free") + coord_cartesian(ylim = c(0, 15))

LM_score<-full_list%>%
        transmute(gene_symbol, end, count_mean = rowMeans(select(., -c(gene_symbol,end)))) %>%
        mutate(norm_score=count_mean/mean(count_mean), .keep="unused")
```
